Smooth and Persistent Forecasts of German GDP\\ Balancing Accuracy and Stability\thanks{We would like Barbara Rossi and the participants at the Vienna Forecasting Workshop 2025 for their comments and suggestions.}

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

For a better understanding, we here briefly substantiate the optimization criterion \ref{critssa}. To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise and $z_t=\sum_{k=-infty}^\infty \gamma_k x_{t-k}=\sum_{k=-infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, which ensures stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)). Assume we want to predict $z_{t+\delta}$, $\delta\geq 0$ and let $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$ designate a causal predictor of length $L$. For simplicity of exposition we may assume $\delta=0$ (nowcast) and we may drop $\delta$ from our notation. Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{k=0}^{L-1}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where 
\[
\mathbf{M}=\left(\begin{array}{ccccccccc}0&0.5&0&0&0&...&0&0&0\\
0.5&0&0.5&0&0&...&0&0&0\\
...&&&&&&&&\\
0&0&0&0&0&...&0.5&0&0.5\\
0&0&0&0&0&...&0&0.5&0
\end{array}\right),
\]
ofs dimension $L\times L$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$ corresponds to the order-one autocovariance of $y_t$, up to scaling by the variance $\sigma^2$ of $\epsilon_t$. We can now rewrite criterion \ref{critssa} as:
\begin{eqnarray}\label{crit1}
\left.\begin{array}{cc}
&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
&\mathbf{b}'\mathbf{b}=l
\end{array}\right\}.
\end{eqnarray}
where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ ensures uniqueness of the SSA solution. 
By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}. \\

>>>>

For a better understanding, we here briefly substantiate the optimization criterion \ref{critssa}. To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise and $z_t=\sum_{k=-infty}^\infty \gamma_k x_{t-k}=\sum_{k=-infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, which ensures stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)). Assume we want to predict $z_{t+\delta}$, $\delta\geq 0$ and let $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$ designate a causal predictor of length $L$. For simplicity of exposition we may assume $\delta=0$ (nowcast) and we may drop $\delta$ from our notation. Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{k=0}^{L-1}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where 
\[
\mathbf{M}=\left(\begin{array}{ccccccccc}0&0.5&0&0&0&...&0&0&0\\
0.5&0&0.5&0&0&...&0&0&0\\
...&&&&&&&&\\
0&0&0&0&0&...&0.5&0&0.5\\
0&0&0&0&0&...&0&0.5&0
\end{array}\right),
\]
ofs dimension $L\times L$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$ corresponds to the order-one autocovariance of $y_t$, up to scaling by the variance $\sigma^2$ of $\epsilon_t$. We can now rewrite criterion \ref{critssa} as:
\begin{eqnarray}\label{crit1}
\left.\begin{array}{cc}
&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
&\mathbf{b}'\mathbf{b}=l
\end{array}\right\},
\end{eqnarray}
where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes, namely to warrant uniqueness of the SSA solution and to ensure that the objective function $\mathbf{b}'\boldsymbol{\gamma}_{\delta}$ is identical to the target correlation, up to an arbitrary scaling term $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$. 
By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}. \\
